---
title: "Workflow Basics Exercises"
author: "Hadley Wickham and Garrett Grolemund and Kim Roth; Edited by Johnathan Clementi"
date: "1/25/2020"
output:
  word_document: default
  pdf_document: default
---


These are somewhat modified exercises from the Workflow Basics Chapter of R For Data Science by Wickham and Grolemund. 

The website where the book appears is (and will always be) __free to use__, and is licensed under the [Creative Commons Attribution-NonCommercial-NoDerivs 3.0](http://creativecommons.org/licenses/by-nc-nd/3.0/us/) License.
```{r setup}
library(tidyverse)
```

1.  Why does this code not work? Write code that works.



    ```{r, error = TRUE}
    my_variable <- 10
    my_variabl
    ```

This code will not run because the 'my_variabl' is not 'my_variable' 
    
```{r}
    my_variable <- 10
    my_variable
```
   
   
   Look carefully! (This may seem like an exercise in pointlessness, but
    training your brain to notice even the tiniest difference will pay off
    when programming.)
    

2.  Tweak each of the following R commands so that they run correctly:
 
```{r, eval = FALSE}
    library(tidyverse)

    ggplot(datta = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy))
    
    fiter(mpg, cyl = 8)
    filter(Diamonds, carat > 3)
```
 
  What I fixed:
  1. 'data' = (not 'datta')
  2. 'filter' (not 'fiter')
  3. '==' comparison operator (not '=' assignment operator)
  4. 'diamonds' (not 'Diamonds')
  
  
```{r}
    library(tidyverse)

    ggplot(data = mpg) + 
      geom_point(mapping = aes(x = displ, y = hwy))
    
    filter(mpg, cyl == 8)
    filter(diamonds, carat > 3)
```
  
   


3.  Press Alt + Shift + K(pc) or Option+Shift_+K(mac). What does it do?

Shows keyboard shortcuts! That's awesome! I will use that SO much


