---
title: "Factors and Dates Assignment"
author: "Kim Roth & Johnathan Clementi"
date: "4/14/2020"
output:
  word_document: default
  pdf_document: default
---


These are exercises based on the exercises in R for Data Science by Wickham and Grolemund.


The website where the book appears is (and will always be) __free to use__, and is licensed under the [Creative Commons Attribution-NonCommercial-NoDerivs 3.0](http://creativecommons.org/licenses/by-nc-nd/3.0/us/) License.



```{r}
library(tidyverse)
gss_cat #our data set for the first part. See the book for details
gssCopy = gss_cat # Make copy for edititing
```

#Factors

1.  Explore the distribution of `partyid` (reported political party preference). What is the default order of factors? Make a plot with reordered factors.  Describe what your best graph tells you about `partid`.
1A. 
Default order: No answer, Don't Know, Other party, Strong republican, Not str republican, Ind, near rep,
Independent, Ind, near dem, Not str democrat, Strong democrat


```{r}
count(gss_cat, partyid)

gssCopy = mutate(gssCopy, partyid = factor(partyid, levels = c("Strong democrat", "Not str democrat", "Ind,near dem", "Independent", "Ind,near rep", "Not str republican", "Strong republican", "Other party", "Don't know", "No answer")))

count(gssCopy, partyid)
```


Load ColorBrewer
```{r}
### Load the package or install if not present
if (!require("RColorBrewer")) {
install.packages("RColorBrewer")
library(RColorBrewer)
}
```

1A(pt2). This figure shows the distribution of political party affiliation amoung people sampled in the GSS. 
There seem to be more people that identify as Democrat (and not strong democrat) than identify as Republican
(or not strong republican). I filtered out 3rd party, Don't know, and didn't answer data to get the color ramp
to fit well with the Democrat - Republican (Left - Right) spread.
```{r}
ggplot(data = filter(gssCopy, partyid != "No answer" & partyid != "Don't know" & partyid != "Other party")) +
  geom_bar(mapping = aes(x = partyid, fill = partyid, color = 'black')) +
  scale_fill_brewer(palette = "RdBu", direction = -1) +
  labs(x = "Party Affiliation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
    
    


2.What is the most common `denom` in this survey? What's the most common `marital`?
2A. Most common `denom` is "Not applicable"; most common `marital` is "Married"
```{r}
denomCount = count(gssCopy, denom)
denomCount = arrange(denomCount, desc(n))
denomCount

maritalCount = count(gssCopy, marital)
maritalCount = arrange(maritalCount, desc(n))
maritalCount
```


    
    
3.  Examine the distribution of `age`. Are there any surprises about the distribution? Would you prefer to use the mean or the median to summarize `age`?
3A. The distribution is strange - there seem to be short boom and busts between age classes, that
almost make the figure look like a hand with too many fingers. I'm not sure what would cause this.
I'd prefer the median over the mean: the 1st and 3rd Standard deviations are equal distance from 
the median (13), whereas they are different values when using the mean. Additionally, the median 
is a whole number (age is traditionally measured in whole years).

```{r}
ggplot(data = gssCopy) +
  geom_histogram(mapping = aes(x = age))

summary(gssCopy$age, na.rm = TRUE)
```

4. For `rincome`, what order are the factors in? Is this a useful order?
4A. Order of factors: No answer, Don't know, Refused, $25000 or more, $20000 - 24999, $15000 - 19999,
$10000 - 14999, $8000 to 9999, $7000 to 7999, $6000 to 6999, $5000 to 5999, $4000 to 4999, $3000 to 3999,
$1000 to 2999, Lt $1000, Not applicable

The order is useful for all of the numeric ranges. However, the first three factors: "No answer, Don't know, Refused" should be moved to the bottom of the factor list.
```{r}
count(gssCopy, rincome)
```

5.  For each factor in `gss_cat` identify whether the order of the levels is arbitrary or principled.
5A.
marital: arbitrary
race: principled
rincome: principled
partyid: principled
relig: arbitrary
denom: principled

```{r, eval = FALSE}
count(gssCopy, marital)
count(gssCopy, race)
count(gssCopy, rincome)
count(gssCopy, partyid)
count(gssCopy, relig)
count(gssCopy, denom)
```
    
6.  How could you collapse `partyid` into a small set of categories? Collapse `partyid` into a small set of categories.
6A. I would collapse into 4 categories: Democrat, Independent, Republican, Other
```{r}
gssCopy = gssCopy %>%
  mutate(partyid = fct_collapse(partyid,
         Democrat = c("Strong democrat", "Not str democrat"),
         Independent = c("Ind,near dem", "Independent", "Ind,near rep"),
         Republican = c("Strong republican", "Not str republican"),
         Other = c("No answer", "Don't know", "Other party")
  ))

count(gssCopy, partyid)
```


7. Use fct_lump to collapse marital. What got lumped together and why?
7A.`No answer`, `Seperated`, and `Widowed` were all lumped together because they were the three factors
with the lowest counts (least common)
```{r}
count(gssCopy, marital)

gssCopy = gssCopy %>%
  mutate(marital = fct_lump(marital))

count(gssCopy, marital)
```


#Dates

```{r}
library(lubridate) #unlike most of the tidyverse libraries, this must be called seperately.
```
8.  What happens if you parse a string that contains invalid dates?
8A. Lubridate throws an `failed to parse` error
    ```{r, eval = FALSE}
    ymd(c("2020-01-10", "2020-13-41", "date"))
    ```



9.  Use the appropriate lubridate function to parse each of the following dates or date-times:

    ```{r}
    d1 <- "Apr 14, 2020"
    d2 <- "2001-21-02"
    d3 <- c("14-01-2021", "15-02-2021")
    d4 <- "01/31/2017 08:01 EST"
    d5 <- "2022-12-31 23:59:59"
    ```
9A.
```{r}
d1=mdy(d1)
d2=ydm(d2)
d3=dmy(d3)
d4=mdy_hm(d4, tz = "America/New_York")
d5=ymd_hms(d5)
d1
d2
d3
d4
d5
```


10. Figure out how old you are in years exactly using R.
10A.
```{r}
JMCBirthday = ymd_hm("1996-05-12 16:11 EST", tz = "America/New_York")
JMCAge = now() - JMCBirthday
as.duration(JMCAge)
```


11.  Why is there `months()` but no `dmonths()`?
11A. The smallest unit of time in a duration is the second. For each d*unit (second, minute, hour, etc),
there is a known number of seconds in each of those units. Because months are variable in their length,
the duration in seconds for each month is not constant.


12.  Explain `days(overnight * 1)` to someone who has just started 
  learning R. How does it work?
12A. To account for periods that exist in multiple days (i.e. a person begins traveling on Sunday and
arrives at their destination on Monday), we must tell R which direction to calculate the time difference,
so it does not result in a negative time value (when departure time is greater than arrival time).


13.  Create a vector of dates giving the first day of every month in the current year.
13A. 
```{r}
firstDay = floor_date(today(), "year")
firstDayList = seq.Date(firstDay, (firstDay + years(1)), "month")
firstDayList
```



14.  Why can't `(today() %--% (today() + years(1)) / months(1))` work?
14A. This is only counting the number of first day of every month (12), it does not create the date
and append it to a list. 

15. Give at least one example that makes working with time zones difficult.
15A. There are some locations which fall in the same time zone but have different policies on observing day-light-savings time, time calculations may need extra care in those cases.
