---
output:
  word_document: default
  pdf_document: default
  html_document: default
---
title: "Model Basics Exercises"
author: "Kim Roth & Johnathan Clementi"
date: "4/7/2020"
output: word_document
---
These are exercises based on the exercises in R for Data Science by Wickham and Grolemund.


The website where the book appears is (and will always be) __free to use__, and is licensed under the [Creative Commons Attribution-NonCommercial-NoDerivs 3.0](http://creativecommons.org/licenses/by-nc-nd/3.0/us/) License.



We will use the data set fueleconomy:vehicles
```{r}
library(tidyverse)
library(fueleconomy)
vehicles<-fueleconomy::vehicles
```


Data for this file.
```{r}
vehicles
```


1. Make a linear model of cty as prediced by displ. Give the equation of the line.
1A.
```{r}
ctyMpgModel = lm(cty~displ, data = vehicles)
ctyMpgModel
```

```{r}
ctyMpgModYInt = 25.812
ctyMpgModCoef = -2.518
```


2. Graph the linear model from 1 using ggplot. How does the fit look?
2A. Without further analysis, the line of fit follows the data well. Based on this figure, smaller 
displacement values are less predictive of city mpg than vehicles with larger displacement.
```{r}
ggplot(data = vehicles) + 
  geom_point(mapping = aes(x = displ, y = cty)) +
  geom_smooth(mapping = aes(x = displ, y = cty), method = "lm", se=0)
```


3. Look at the first residual graph of the linear model. How does it look?
3A. The dataset has a good constant spread and while the average value closely aligns with a '0' residual
in the middle of the data, the upper and lower bounds do not fit as well. Additionally, the spread of 
residuals at the upper bound means that as displacement increase, the model is less able to predict the
city mpg.
```{r}
plot(ctyMpgModel)
```


4.  Instead of using `lm()` to fit a straight line, you can use `loess()` 
    to fit a smooth curve. Repeat the model from 1. using
    `loess()` instead of `lm()`. How does the result compare to 
    `geom_smooth()`? What are the advantages and disadvantages to using a curve versus a line here?
4A. The curved line presents a different message to the reader than a straight line does. For instance,
looking at the linear model figure, I would identify that there is more variation of city mileage in cars
with less displacement. Whereas looking at the locally estimated scatterplot smoothing figure, 
I would more quickly highlight that cars with smaller displacement have better city mpg ratings. 
```{r}
ctyMpgCurveModel = loess(cty~displ, data = vehicles)
ctyMpgCurveModel
```

```{r}
ggplot(data = vehicles) + 
  geom_point(mapping = aes(x = displ, y = cty)) +
  geom_smooth(mapping = aes(x = displ, y = cty), method = "loess", se=0)
```

    
In case you want to know more LOESS stands for, locally estimated scatterplot smoothing,it essentially puts together a bunch of linear for parts of the data, see https://simplystatistics.org/2014/02/13/loess-explained-in-a-gif/

5. Predict the cty for displ of 1.9. 
5A.
```{r}
predictedDispl = (ctyMpgModCoef * 1.9) + ctyMpgModYInt
predictedDispl
```


6. Graph a scatterplot of cty vs. drive. Do you think a model is a good idea?
6A.
```{r}
ggplot(data = vehicles) +
  geom_point(mapping = aes(x = drive, y = cty))
```


7. Make a linear model for cty as predicted by drive.
```{r}
ctyDriveModel = lm(cty~drive, data = vehicles)
ctyDriveModel
```
```{r}
ctyDriveFWDCoef = 4.1005
ctyDriveYInt = 16.8915
```


8. Look at the first residual graph of the linear model from 7. How does it look?
8A. Judging by the average line vs the 0 residual line, the drive variable is a good predictor of
city mpg. 
```{r}
plot(ctyDriveModel)
```


9. Predict the cty for the drive of 2-Wheel Drive.
9A.
```{r}
predictedDrive = (ctyDriveFWDCoef * 1) + ctyDriveYInt
predictedDrive
```


10.  Make scatterplots for the quantitative variables versus cty. Which looks like it is going to make the best linear model. Why?
10A. City mpg and hwy mpg plotted against eachother seem like the best for a linear model. The spread of data
indicate that generally as cty increases, so does hwy.
```{r}
ggplot(data = vehicles) + 
  geom_point(mapping = aes(x = year, y = cty))

ggplot(data = vehicles) + 
  geom_point(mapping = aes(x = cyl, y = cty))

ggplot(data = vehicles) + 
  geom_point(mapping = aes(x = displ, y = cty))

ggplot(data = vehicles) + 
  geom_point(mapping = aes(x = hwy, y = cty))
```


11. Make a linear model based on either your best pick in 10  or the second best pick if displ was best pick and graph it.
11A. 
```{r}
ggplot(data = vehicles) + 
  geom_point(mapping = aes(x = hwy, y = cty)) +
  geom_smooth(mapping = aes(x = hwy, y = cty), method = "lm", se = 0)
```


12.Which of the models predicting hwy would you prefer to use and why?
12A. I would use the locally estimated scatterplot smoothing method rather than the linear model method
as it more closely follows the spread of data.
```{r}
hwyCtyLmModel = lm(cty~hwy, data = vehicles)
hwyCtyLmModel

hwyCtyLoessModel = loess(cty~hwy, data = vehicles)
hwyCtyLoessModel
```

```{r}
plot(hwyCtyLmModel)

ggplot(data = vehicles) + 
  geom_point(mapping = aes(x = hwy, y = cty)) +
  geom_smooth(mapping = aes(x = hwy, y = cty), method = "lm", se = 0)
```


```{r}
plot(hwyCtyLoessModel)

ggplot(data = vehicles) + 
  geom_point(mapping = aes(x = hwy, y = cty)) +
  geom_smooth(mapping = aes(x = hwy, y = cty), method = "loess", se = 0)
```


13.Choose a new pair of variables to make a linear model. The response variable must be quanitiative. Graph your model and give an equation.
13A.y = (0.1391 * x) - 254.4613
```{r}
hwyYearModel = lm(hwy~year, data = vehicles)
hwyYearModel

hwyYearFunc = function(x){
  (0.1391 * x) - 254.4613
}
```


```{r}
plot(hwyYearModel)

ggplot(data = vehicles) + 
  geom_point(mapping = aes(x = year, y = hwy)) +
  geom_smooth(mapping = aes(x = year, y = hwy), method = "lm", se = 0)
```


14. How is the fit?
14A. Considering the data presented, the fit is good - as we progress in time, generally cars are becoming
more fuel efficient. 

15. Make a prediction with your model.
15A.
```{r}
hwyYearFunc(2030)
```

