---
title: "Modeling Summer Read"
author: "Kim Roth"
date: "8/7/2019"
output: word_document
---

Today we will model the Summer Reading Program Data.
```{r, results="hide", warning=FALSE, message=FALSE, error=FALSE}
library(tidyverse) #loads ggplot and the others

library(readr)#for reading in csv files. .
```

Data import. Make sure you use the new, cleaner data for this file. Ignore the lengthy error messages. It does fine in the end.
```{r}
SummerReadClean <- read_csv("C:/Users/clemenj/Google Drive/Grad School/Juniata_DataScience/DS500/Week11/SummerReadClean.csv")  
```

```{r}
ggplot(data=SummerReadClean)+geom_point(mapping=aes(x=BooksReadSummer, y=PercentReadSummer))
```


Predicting PercentReadSummer
```{r}
model1=lm(PercentReadSummer~BooksReadSummer, data=SummerReadClean)
model1
```
x

Looking at model
```{r}
ggplot(data=SummerReadClean)+geom_point(mapping=aes(x=BooksReadSummer, y=PercentReadSummer))+geom_abline(mapping=aes(intercept=model1$coef[1], slope=model1$coef[2])) #this is the books method, par stands for parameter
ggplot(data=SummerReadClean)+geom_point(mapping=aes(x=BooksReadSummer, y=PercentReadSummer))+geom_smooth(mapping=aes(x=BooksReadSummer, y=PercentReadSummer), method="lm",se=0) #my prefered method
```
Residual plots. We want the first one. Ignore the others
```{r}
plot(model1)
```
With a categorical variable
```{r}
count(SummerReadClean, Gender)
ggplot(data=SummerReadClean)+geom_point(mapping=aes(y=PercentReadSummer, x=Gender))
```
Modeling with categories
```{r}
model2=lm(PercentReadSummer~Gender+BooksReadSummer, data=SummerReadClean)
model2
summary(model2)
```
Making predictions. Right now the hard way. Book way some other time.
```{r}
predictModel2 <-
  data_frame(
    Gender = c("Female", "Male", "Trans/NonBinary", NA),
    PercentReadSummer = c(42.38, 52.38-12.25, 42.38-12.04, NA)
  )
```

```{r}
ggplot(data=SummerReadClean)+geom_point(mapping=aes(x=Gender, y=PercentReadSummer))+geom_point(data=predictModel2, mapping=aes(x=Gender, y=PercentReadSummer), color="blue",size=5)
```
Residuals, remember only the first graph.
```{r}
plot(model2)
```

Now it is your turn. Predict PercentReadSemester. What relationships are you curious about? Make a one and two  predictor models.  
