---
output:
  word_document: default
  html_document: default
---
title: "Strings  Exercises"
author: "Kim Roth & Johnathan Clementi"
date: "04/28/20"
output: word_document
---
These are exercises based on  R for Data Science by Wickham and Grolemund. 


The website where the book appears is (and will always be) __free to use__, and is licensed under the [Creative Commons Attribution-NonCommercial-NoDerivs 3.0](http://creativecommons.org/licenses/by-nc-nd/3.0/us/) License.

While not required I reccomend the regex crossword for practice
https://regexcrossword.com

```{r setup, message = FALSE}
library(tidyverse)
```

1.  In code that doesn't use stringr, you'll often see `paste()` and `paste0()`.
    What's the difference between the two functions? What stringr function are
    they equivalent to? How do the functions differ in their handling of 
    `NA`?
    
1A. `paste()` is different from `paste0()` in that they concatenate strings, but `paste()` allows for a seperator
arguments to be included. They are equivilent to the `str_c()` function of stringr. Both `paste()` and `paste0()`
deal with NA as a string, but `str_c` deals with NA as a missing value. 


2.  Use  `str_sub()` to extract the second to last character from a string.
```{r}
x <- c("Apple", "Banana", "Pear")

str_sub(x, -2, -2)
```



3.  What is the difference between `n`, `\n`, `\\n`, `\\\n`?
`n` matches the character "n"
`\n` matches a new line if put into a regex function
`\\n` can be passed as a string to a regex function and will match a new line - `\\` will match a \ in a string
`\\\n` if the first \ is an escape character for the second \, then the regex will match a \ before a new line. 



 For 4-13 Given the corpus of common words in `stringr::words`, create regular
    expressions that find all words that:
```{r}
head(words,5)
```
    
4. Start with "q".
```{r}
str_subset(words, "^q")
```

5. End with "y"
```{r}
str_subset(words, "y$")
```

    
6. Have five letters or more.
```{r}
str_subset(words, "[:alpha:]{5,}")
```


7. Start with a consontants.(Hint: thinking about matching "not"-vowels.)
```{r}
str_subset(words, "^[b-df-hj-np-tv-z]")
```
```{r}
str_view(words,"^[^aeiou]", match=TRUE)
```



8. That only contain vowels. 
```{r}
str_subset(words, "^[aeiou]+$")
```


9. End with `ing`, but not with `ring`.
```{r}
str_subset(words, "[^r]ing$")
```
    

    
10. End with `ed` or `ly`.
```{r}
str_subset(words, "(ed|ly)$")
```


11. Start with two vowels.
```{r}
str_subset(words, "^[aeiou]{2}")
```

    
12. Have three or more consontant in a row.
```{r}
str_subset(words, "[b-df-hj-np-tv-z]{3,}")
```


13. Have two or more consonant-vowel pairs in a row.
```{r}
str_subset(words, "([b-df-hj-np-tv-z][aeiou]){2,}")
```


14. Replace spaces in a string with _
```{r}
cars = tibble::rownames_to_column(mtcars,"Name")
cars = head(cars, 35)
```

```{r}
str_replace(cars$Name,"\\s","_")
```


