title: "Strings  Exercises"
author: "Kim Roth"
date: "12/12/2019"
output: word_document
---
These are exercises based on  R for Data Science by Wickham and Grolemund. 


The website where the book appears is (and will always be) __free to use__, and is licensed under the [Creative Commons Attribution-NonCommercial-NoDerivs 3.0](http://creativecommons.org/licenses/by-nc-nd/3.0/us/) License.

While not required I reccomend the regex crossword for practice
https://regexcrossword.com

```{r setup, message = FALSE}
library(tidyverse)
```


1.  In code that doesn't use stringr, you'll often see `paste()` and `paste0()`.
    What's the difference between the two functions? What stringr function are
    they equivalent to? How do the functions differ in their handling of 
    `NA`?
1A.


paste() and paste0() treat NA like a text string. str_c treats it like NA.

2.  Use  `str_sub()` to extract the second to last character from a string.



3.  What is the difference between `n`, `\n`, `\\n`, `\\\n`? 





 For 4-13 Given the corpus of common words in `stringr::words`, create regular
    expressions that find all words that:
```{r}
head(words,5)
```
    
4. Start with "q".

    
5. End with "y"

    
  
    
6. Have five letters or more.

7. Start with a consontants.(Hint: thinking about matching 
       "not"-vowels.)


8. That only contain vowels. 



9. End with `ing`, but not with `ring`.
    

    
10. End with `ed` or `ly`.



11. Start with two vowels.

    
12. Have three or more consontant in a row.
 
13. Have two or more consonant-vowel pairs in a row.



14.   Replace spaces in a string with _


