---
title: "Introduction to Graphing with ggplot2"
author: "Kim Roth"
date: "9/30/2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

**Introduction**

This is an RMarkdown document. It allows you type text amid your r commands without having to comment. It can produce professional level documents in html, word, or pdf. 

This is an R chunk. It is like an embedded .R file. Hit the play button to evaluate. You can also hit shift+control+enter for PCs or shift+command+enter for macs.
```{r}
3+4
sample=c(1,2,3,4)
sample
```

Note the output is displayed in two places, in the markdown file and in the Console, althogh RStudio sometimes likes to hide the console.

You can expand or contract your output by clicking on the ^^ symbol next to the x in the white output area.

Two things to remember about programming in R.

1. The computer will do exactly what you tell it to do. 
2. Copy, paste, and edit is your good friend.

Albert and Chester like to say instead
1. Computers are stupid.
2. Don't code from scratch.

Now like many programming environments we have to load in a few things to use them here. In R we do this with the library()  command. Technically you have to install the packages before you load them, but we've done that already.


```{r}
library(tidyverse) #loads ggplot and the others

simple_ex <-
  data_frame(
    A = c(1980, 1990, 2000, 2010),
    B = c(1, 2, 3, 4),
    C = c(3, 2, 1, 2),
    D = c("low", "low", "high", "high")
  ) #the example from class
simple_ex #looking at the example data
```

**Graphs from simple_ex**
Now let's make a graph starting with a set of axes, called the base layer
```{r fig.height=2, fig.width=5}
ggplot(data = simple_ex, mapping = aes(x = A, y = B)) #base layer, there is no graph here yet, just an x and y axis
```



Let's pick that commmand apart a bit.

ggplot-calls ggplot
data= tells what data set
mapping= what variables we will graph (optional, can be omitted)
aes is short for aesthetic

Now let's make a scatterplot. A graph is a mapping of data variables to aestheic attributes to geometric object. The graph type is a geometric object. Scatter plots are geom_point. To add things to a graph use +. For formatting reasons it is best to end lines with +.

There are two possible places to put the aesthetic information when you have a geom. Either in the ggplot part of the geom part. We will usually follow R for Data Science. Most of the time they are equivalent and you may use either. See below.
```{r}
ggplot(data = simple_ex) +geom_point(mapping=aes(x = A, y = B)) #R for Data Science Style
ggplot(data = simple_ex, mapping = aes(x = A, y = B)) +geom_point() #Another Style

```
And there is our first graph.

Now to color by D.

```{r}
ggplot(data = simple_ex) +geom_point(mapping=aes(x=A, y=B, color = D))
```

Now to do size by C.
```{r}
ggplot(data = simple_ex) + 
  geom_point(mapping = aes(x = A, y = B, size = C))
```

Now color and size.
```{r}
ggplot(data = simple_ex) +geom_point(mapping = aes(x = A, y = B,size=C, color = D))
```

Line graphs use the geometry geom_line
```{r}
ggplot(data = simple_ex) + 
  geom_line(mapping = aes(x = A, y = B))
```

Now coloring the line according to D and points blue of size 4. The coloring and sizing of the points goes inside the geom_point() command because it effects the points, but outside the aes() part becuase it is not messing with the variables.

What will happen with the middle of the line?
```{r}
ggplot(data = simple_ex) + 
  geom_line(mapping = aes(x = A, y = B,color = D)) +
  geom_point(mapping=aes(x = A, y = B),color = "blue", size = 4)
```

When using more than one geom it is easier to put the aesthetics that are in common in the ggplot part.
```{r}
ggplot(data = simple_ex, mapping = aes(x = A, y = B,color = D)) + 
  geom_line(mapping=aes(color = D)) +
  geom_point(color = "blue", size = 4) #this gives the same graph.
```
**Notes**

Common graph types

- Scatterplot: `geom_point()`
- Line graph: `geom_line()`
- Histogram: `geom_histogram()`
- Boxplot: `geom_boxplot()`
- Bar graph: `geom_bar()`
- Density plots: `geom_density()'


You can find more on the ggplots cheatsheet.

Now it is time for you to make your own graphs. Proceed to the file OKCupidMoreGgplot.Rmd
