---
title: "Tidyr And Functions Exercises"
author: "Kim Roth & Johnathan Clementi"
date: "3/27/2020"
output:
  word_document: default
  pdf_document: default
---


These are exercises based on the exercises in R for Data Science by Wickham and Grolemund.


The website where the book appears is (and will always be) __free to use__, and is licensed under the [Creative Commons Attribution-NonCommercial-NoDerivs 3.0](http://creativecommons.org/licenses/by-nc-nd/3.0/us/) License.



We will use the data set fueleconomy:vehicles
```{r}
library(tidyverse)
library(fueleconomy)
vehicles<-fueleconomy::vehicles
```

Data for this file.
```{r}
vehicles
```
A subset for parts of the homework.
```{r}
hybridSubset=filter(vehicles, fuel=="Regular Gas and Electricity")
hybridSubset=select(hybridSubset, make, model, year, hwy)
hybridSubset
```


1.  Why are `gather()` and `spread()` not perfectly symmetrical?  
    Carefully consider the following example:
    
```{r, eval = FALSE}

hybridSubset
    
hybridSubset %>% 
  spread(year, hwy) %>% 
  gather("year", "hwy", `2012`:`2014`)
```

1A. `Gather` and `Spread` are not totally symmetric if there are instances of non-existant observations for all categories within all variables
in the starting data, known as implicit NA. For instance, the original data here does not have an observation for the Ford C-MAX Energi Plug-in in 2012. 
However, after spreading and then gathering, the observation is created and only has NA for the hwy reading. 


2. Make the command from 1 give back the original data. You can use spread and gather or pivot_wider and pivot_longer.  
2A.
```{r}
hybridSubset %>% 
  spread(year, hwy) %>% 
  gather("year", "hwy", `2012`:`2014`) %>%
  filter(!is.na(hwy))
```

3. Why does spread or pivot_wider on this tibble fail? Why doesn't it fail on the full `hybridSubset`?

```{r, eval=FALSE}
hybridSubset2 = select(hybridSubset, year, make, hwy)

hybridSubset2 %>%
  spread(year, hwy)
```
3A. Spread or pivot_wider fails to spread this tibble because it cannot create a unique key for each new observation as it spreads


4. Make a graph of the average hwy, cty, and displ with the original data set `vehicles`. Use the example in the videos to help you.
4A.
```{r}
mpgDispl = select(vehicles, make, model, year, hwy, cty, displ)
  
mpgDispl = mpgDispl %>%  
  pivot_longer(c("hwy", "cty", "displ"), names_to = "mpg", values_to = "Average")

ggplot(data = mpgDispl) +
  geom_boxplot(aes(x = mpg, y = Average)) +
  labs(x = "", y = "Measurement (miles per gallon)")
```



5. Use separate to split the displ variable into parts named integer and decimal using the data set `vehicles` (Yes this is a silly exercise, but the data set did not have a better option.)
5A.
```{r}
vehicles = separate(vehicles, displ, into = c("displInt", "displDec"))
```


6. Use unite to combine the variables make and model into one column called car in the original data set, `vehicles`.
6A.
```{r}
vehicles = unite(vehicles, car, make, model)
```


7. Give an example of an explict NA from `vehicles`. Give an example of a possible implicit NA from `vehicles `.
7A. Explicit NA: hwy, cty, and displ measurements for electric cars
    Implicit NA: observations for cars that were not made during certain years in the study 
    (i.e. car wasn't created until certian year, or car was discontinued after a certain year)


8. Give an example of a data set (other than the one in the book) where using fill in data set is a good idea. Give an example of a data set where using fill is a bad idea.
8A. A good candidate data set for using fill on is one that has many redundant observations. Using fill on a data set with many unique observations in succession
could introduce a lot of bad data.

 
9.  Why is `TRUE` not a parameter to `rescale01()` from the reading? What would happen if
    `x` contained a single missing value, and `na.rm` was `FALSE`? 
9A. The function rescales a series of numbers to a scale from 0 to 1. The parameters being passed to the function are numeric, not boolean (which `TRUE` is).
    The missing value would be ignored by the function.
    
```{r}
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
rescale01(c(0, NA, 10))
```
    

10. Write a function that calculates the product of two numbers. Test your function on some values.
10A.
```{r}
prdct = function(x, y){
  x * y
}

prdct(4,4)
prdct(5,5)
prdct(6,6)
```


11.  Practice turning the following code snippets into functions. Think about 
    what each function does. What would you call it? How many arguments does it
    need? 

  
    `sum(x>10,na.rm=TRUE)/sum(!is.na(x))`
    
   `median(x, na.rm = TRUE)`
    
  `mean(x)/ sd(x)`
11A.
```{r}
#This function takes a vector and calculates the sum of all numbers greater than 10 divided by the sum of all of the numbers in the vector
greaterThanTen = function(x){
  sum(x>10, na.rm = TRUE)/sum(!is.na(x))
}

# This function calculates the median of a set of numbers (provided in the form of a vector)
calculateMedian = function(x){
  median(x, na.rm = TRUE)
}

# If the coefficient of variation is the standard deviation over the mean, then this function is the inverse coefficient of variation
inverseCV = function(x){
  mean(x) / sd(x)
}

```


Here is some data to test it on.
```{r}
test1=c(1,2,3,4)
test2=c(0,5,10,15,20)
test3=c(-3,0,3,6,9,12,15,18,21)
test4=c(1,0,100, NA, -1000)

greaterThanTen(test1)
greaterThanTen(test2)


calculateMedian(test1)
calculateMedian(test2)

inverseCV(test1)
inverseCV(test2)
```


12.  Write a function that takes in a vector and outputs the percentage of the vector that is NA.
12A.
```{r}
pctNA = function(x, count = 0){
  for(y in x) {
    if(is.na(y)) {
      count = count + 1
    } else {}
  }
  count/(length(x))
}
```

```{r}
pctNA(test1)
```



13.  Write `same_mean()`, a function that takes two vectors and returns if they have the same mean or not.
13A.
```{r}
same_mean = function(x, y){
  if(mean(x) == mean(y)) {
    return(TRUE)
  } else {
    return(FALSE)
  }
}
```

```{r}
same_mean(test1, test2)
same_mean(test1, test1)
```



14.  What do the following functions do?
14A. path = the working directory of the current R script
     combine = concatenate x and y strings without space between them
    
    ```{r}
    path <- function() getwd()
    combine <- function(x,y) paste0(x, y)
    ```


15. Read the source code for each of the following three functions, puzzle out
    what they do, and then brainstorm better names.
 
    ```{r}
    f1 <- function(string, prefix) {
      substr(string, 1, nchar(prefix)) == prefix
    }
    f2 <- function(x) {
      if (length(x) <= 1) return(NULL)
      x[-length(x)]
    }
    f3 <- function(x, y) {
      rep(y, length.out = length(x))
    }
    ```
15A.
f1 determines if `string` has a `prefix` at the start of the character vector. It could be renamed testForPrefix

f2 removes the last element of a vector. It could be renamed rmvLstElement

f3 replicates or truncates the second vector (y) so that it is the same length as the first vector (x). It could be renamed lengthAdjustor 

