---
title: "R For Data Science Workflow:Scripts Code"
author: "Wickham and Grolemund"
date: "6/25/2019"
output: word_document
---

This is most of the code as it appears in R for Data Science by Wickham and Grolemund. While I have added a few comments about what appears, no changes have been made.


The website where the book appears is (and will always be) __free to use__, and is licensed under the [Creative Commons Attribution-NonCommercial-NoDerivs 3.0](http://creativecommons.org/licenses/by-nc-nd/3.0/us/) License.
# Data import

## Introduction

### Prerequisites

```{r setup, message = FALSE}
library(tidyverse)
```

## Getting started


```{r, message = TRUE}
heights <- read_csv("data/heights.csv")
```


```{r}
read_csv("a,b,c
1,2,3
4,5,6")
```


    
    ```{r}
    read_csv("The first line of metadata
      The second line of metadata
      x,y,z
      1,2,3", skip = 2)
    
    read_csv("# A comment I want to skip
      x,y,z
      1,2,3", comment = "#")
    ```

    ```{r}
    read_csv("1,2,3\n4,5,6", col_names = FALSE)
    ```
  
    
    ```{r}
    read_csv("1,2,3\n4,5,6", col_names = c("x", "y", "z"))
    ```


```{r}
read_csv("a,b,c\n1,2,.", na = ".")
```


### Compared to base R


## Parsing a vector


```{r}
str(parse_logical(c("TRUE", "FALSE", "NA")))
str(parse_integer(c("1", "2", "3")))
str(parse_date(c("2010-01-01", "1979-10-14")))
```


```{r}
parse_integer(c("1", "231", ".", "456"), na = ".")
```

```{r}
x <- parse_integer(c("123", "345", "abc", "123.45"))
```


```{r}
x
```


```{r}
problems(x)
```


### Numbers

```{r}
parse_double("1.23")
parse_double("1,23", locale = locale(decimal_mark = ","))
```



```{r}
parse_number("$100")
parse_number("20%")
parse_number("It cost $123.45")
```

```{r}
# Used in America
parse_number("$123,456,789")

# Used in many parts of Europe
parse_number("123.456.789", locale = locale(grouping_mark = "."))

# Used in Switzerland
parse_number("123'456'789", locale = locale(grouping_mark = "'"))
```

### Strings {#readr-strings}


```{r}
charToRaw("Hadley")
```


```{r}
x1 <- "El Ni\xf1o was particularly bad this year"
x2 <- "\x82\xb1\x82\xf1\x82\xc9\x82\xbf\x82\xcd"

x1
x2
```


```{r}
parse_character(x1, locale = locale(encoding = "Latin1"))
parse_character(x2, locale = locale(encoding = "Shift-JIS"))
```


```{r}
guess_encoding(charToRaw(x1))
guess_encoding(charToRaw(x2))
```

### Factors {#readr-factors}


```{r}
fruit <- c("apple", "banana")
parse_factor(c("apple", "banana", "bananana"), levels = fruit)
```

### Dates, date-times, and times {#readr-datetimes}


    
    ```{r}
    parse_datetime("2010-10-01T2010")
    # If time is omitted, it will be set to midnight
    parse_datetime("20101010")
    ```
  
    
    ```{r}
    parse_date("2010-10-01")
    ```

  
    ```{r}
    library(hms)
    parse_time("01:10 am")
    parse_time("20:10:01")
    ```
  
```{r}
parse_date("01/02/15", "%m/%d/%y")
parse_date("01/02/15", "%d/%m/%y")
parse_date("01/02/15", "%y/%m/%d")
```



```{r}
parse_date("1 janvier 2015", "%d %B %Y", locale = locale("fr"))
```



## Parsing a file


### Strategy


```{r}
guess_parser("2010-10-01")
guess_parser("15:01")
guess_parser(c("TRUE", "FALSE"))
guess_parser(c("1", "5", "9"))
guess_parser(c("12,352,561"))

str(parse_guess("2010-10-10"))
```


### Problems

```{r}
challenge <- read_csv(readr_example("challenge.csv"))
```


```{r}
problems(challenge)
```



```{r, eval = FALSE}
challenge <- read_csv(
  readr_example("challenge.csv"), 
  col_types = cols(
    x = col_integer(),
    y = col_character()
  )
)
```


```{r}
challenge <- read_csv(
  readr_example("challenge.csv"), 
  col_types = cols(
    x = col_double(),
    y = col_character()
  )
)
```


```{r}
tail(challenge)
```



```{r}
challenge <- read_csv(
  readr_example("challenge.csv"), 
  col_types = cols(
    x = col_double(),
    y = col_date()
  )
)
tail(challenge)
```



### Other strategies


   
    ```{r}
    challenge2 <- read_csv(readr_example("challenge.csv"), guess_max = 1001)
    challenge2
    ```

   
    ```{r}
    challenge2 <- read_csv(readr_example("challenge.csv"), 
      col_types = cols(.default = col_character())
    )
    ```
    
  

    ```{r}
    df <- tribble(
      ~x,  ~y,
      "1", "1.21",
      "2", "2.32",
      "3", "4.56"
    )
    df
    
    # Note the column types
    type_convert(df)
    ```
    

## Writing to a file



```{r, eval = FALSE}
write_csv(challenge, "challenge.csv")
```


```{r, warning = FALSE}
challenge
write_csv(challenge, "challenge-2.csv")
read_csv("challenge-2.csv")
```


```{r}
    write_rds(challenge, "challenge.rds")
    read_rds("challenge.rds")
```
  
    
    ```{r, eval = FALSE}
    library(feather)
    write_feather(challenge, "challenge.feather")
    read_feather("challenge.feather")
    #> # A tibble: 2,000 x 2
    #>       x      y
    #>   <dbl> <date>
    #> 1   404   <NA>
    #> 2  4172   <NA>
    #> 3  3004   <NA>
    #> 4   787   <NA>
    #> 5    37   <NA>
    #> 6  2332   <NA>
    #> # ... with 1,994 more rows
    ```


```{r, include = FALSE}
file.remove("challenge-2.csv")
file.remove("challenge.rds")
```

## Other types of data
