---
title: "Visualization-Which graph and labels"
author: "Kim Roth & Johnathan Clementi"
date: "2/12/2020"
output: html_document
---

Libraries needed to run this document and a bit of data formatting.
```{r}
library("tidyverse") #for ggplot
library("fivethirtyeight") # for data.
# Turn off scientific notation
options(scipen = 99)
```

We will look at the data set bechtel from this article https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/ from FiveThirtyEight. 
Here is a description of some of the graphs.
https://cran.r-project.org/web/packages/fivethirtyeight/vignettes/bechdel.html
Here you can find a description of the variables.
https://cran.r-project.org/web/packages/fivethirtyeight/fivethirtyeight.pdf


The Bechdel test is a set of criterion on female rerpresentation in a movie. To pass the test:

1. Movie must contain at least two named women.
2. At least two women must talk to each other.
3. The women must talk to each other about something other than a man.

So now lets look at the data. The cases here are films.
```{r}
str(bechdel) #list variables
bechdel
```

So say we want to graph pass and fail amounts for the test. That is stored in the variable binary.

1. What type of variable is binary?
1A. Categorical

2. What type of graphs can we use?
2A. One discrete variable graphs:  Bar chart (maybe pie chart)

3. Make a graph for binary. Describe what you see. The description should be a sentence or two about interesting features of the graph.

```{r}

```



Now we are going to label your graph. 
To label a graph use the command +labs(title="Title you want", x="X axis title", y="Y axis title").

For example graphing clean_test which is a categorical variable with reasons for failure:
```{r}
ggplot(data = bechdel) +
  geom_bar(mapping = aes(x = clean_test))+labs(title="Bechdel Test Results with Reason Failed", x="Movie Classification",y="Number of Movies")
```



4. Make a new graph for binary with labels.


5. What sorts of graphs could you make using both binary and clean_test?


6. Make a labeled graph using both binary and clean_test. Describe what you see.


7. Using your previous graph for binary and clean_test, change into a different graph by either changing positon or using `coord_polar()`.


8. What type of variable is budget?



9. What sorts of graphs could you make of budget?


10. Make a labeled graph of budget.Describe what you see.


11. Add a categorical/discrete variable to the budget graph as a fill. Change the color scheme to something other than the default. Make sure to adjust the labels. Describe what you see.


12. Graph both budget and binary together with labels using facets. Describe what you see.



13. What kind of variable is decade code? Note decade code represents the groups of decades the data uses. Movies from 2010 and above are decade 1, movince from 2000 to 2009 are decade 2, movies from 1990 to 1999 are decade 3. Movies before 1990 do not have a decade code.


14. How is R treating it?



You might want to use decade code as a category rather than a number. If you want to use decade code as a category, you will need to tell R it is a cateogry, which for R is called a factor. We will do this using as.factor
```{r}
bechdel=mutate(bechdel, decadecat=as.factor(decade_code)) #defines a new variable decadecat that has decade code as a category. Note could just rename old variable too. This is a dplyr command, we'll talk about that soon
str(bechdel)
```

15. Now make a graph of budget_2013 by decadecat. Describe what you see.
```{r}

```

This document based on:
https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/ 
https://cran.r-project.org/web/packages/fivethirtyeight/vignettes/bechdel.html


