title: "Data Transformation with dplyr Exercises"
author: "Hadley Wickham, Garrett Grolemund, and Johnathan Clementi"
date: "2/23/2020"
output: word_document
---

These are exercises based on the exercises in R for Data Science by Wickham and Grolemund.


The website where the book appears is (and will always be) __free to use__, and is licensed under the [Creative Commons Attribution-NonCommercial-NoDerivs 3.0](http://creativecommons.org/licenses/by-nc-nd/3.0/us/) License.

# Data transformation {#transform}

We will use the data set fueleconomy:vehicles
```{r}
library(tidyverse)
library(fueleconomy)
vehicles<-fueleconomy::vehicles
```

from https://cran.r-project.org/web/packages/fueleconomy/fueleconomy.pdf

vehicles- 5 cat., 4 quant, 1 case id
id Unique EPA identifier
make Manufacturer 
model Model name 
year Model year
class EPA vehicle size class, http://www.fueleconomy.gov/feg/ws/wsData.shtml#VClass trans Transmission
drive Drive train
cyl Number of cylinders
displ Engine displacement, in litres fuel 
Fuel type
hwy Highway fuel economy, in mpg 
cty City fuel economy, in mpg
```{r}
vehicles
```
```{r}
count(vehicles, make)
count(vehicles, model)
count(vehicles, year)
count(vehicles, class)
count(vehicles, trans)
count(vehicles, drive)
count(vehicles, cyl)
count(vehicles, displ)
count(vehicles, fuel)
count(vehicles, hwy)
count(vehicles, cty)
```

### Exercises

1.  Find all vehicles that

    a. were made after 2010
    b. were made by Ford or General Motors
    c.  had 2-Wheel Drive or Front-Wheel Driv
    d. had a manual transmition of any kind. 
    e. had city miles per gallon above the highways miles per gallon.
    f. had highway miles per gallon above 25 and less than 30.
    g. had displacement between 1 and 5 (inclusive)

1A.
```{r}
    #a.
    filter(vehicles, year > 2010)
    
    #b. 
    filter(vehicles, make == 'Ford' | make == 'General Motors')

    #c.
    filter(vehicles, drive == '2-Wheel Drive' | drive == 'Front-Wheel Drive')

    #d.
    filter(vehicles, grepl('Manual', trans))
    
    # Used grepl for regex insteal of typing out each name... which would look like:
    
    filter(vehicles, trans == 'Manual 3-spd' | trans == 'Manual 4-spd' | trans == 'Manual 5-spd' | trans == 'Manual 6-spd' | trans == 'Manual 7-spd' | trans == 'Manual(M7)')
    
    #e.
    filter(vehicles, cty > hwy)
    # Important: still contains NA values...

    #f.
    filter(vehicles, hwy > 25 & hwy < 30)

    #g.
    filter(vehicles, displ >= 1 & displ <= 5)
```


2.  Another useful dplyr filtering helper is `between()`. What does it do?
    Can you use it to simplify the code needed to answer the previous 
    part?
    
2A. The `between()` function filters data vector that fall between the left and right bounds (inclusive). It is useful for simplifying code.
```{r}
filter(vehicles, between(hwy, 26, 29))
filter(vehicles, between(displ, 1, 5))
```



3.  How many cars have a missing `cyl`? What other variables are 
    often missing with cyl? What might these rows represent?
    
3A. There are 58 cars in this list that do not have `cyl` values. These are mainly electric vehicles that do not have traditional combustion chambers (cylinders). However, the Mazda RX-7 and the Ford Ranger Pickups from 1999-2001 are not electric and are just missing the `cyl` values.
```{r}
filter(vehicles, is.na(cyl))
```


    
4. Why is `NA ^ 0` not missing? Why is `NA | TRUE` not missing?
    Why is `FALSE & NA` not missing? Can you figure out the general
    rule?  (`NA * 0` is a tricky counterexample!)

4A. `NA ^ 0` and `NA | TRUE` evaluate as `1` and `TRUE` respectively because they use an type of OR comparison operator, which ignores the presence of an unknown value. `FALSE & NA` evaluates as `FALSE`. I think the general rule is that when there is a comparison operator which allows us to 'ignore' the NA value.

5.  How could you use `arrange()` to sort all missing values of cyl to the start?
    (Hint: use `is.na()`).
    
5A. 
```{r}
arrange(vehicles, !is.na(cyl), cyl)
```


    
6.  Sort `vehicles` to find the vehicles with the largest city miles per gallon. Also find the vehicles that had the lowest highway miles per gallon.

6A. 
Highest CTY mpg: 138
Lowest HWY mpg: 9
```{r}
arrange(vehicles, desc(cty))

arrange(vehicles, hwy)
```



7. Sort `vehicles` to find the vehicles with the largest number of cylnders.

7A. 
Largest number of cylinders: 16
```{r}
arrange(vehicles, desc(cyl))
```


8.  Which vehicles have the largest displacement? Which have the smallest?

8A. 
Largest displacement: 8.4
Smallest displacement: 0.0? or 1.0
```{r}
arrange(vehicles, desc(displ))
arrange(vehicles, displ)
```


    
9.  Select all quantitative variables from `vehicles`.

9A. cyl, displ, hwy, cty
```{r}
select(vehicles, cyl, displ, hwy, cty)
```


10. What does the `one_of()` function do? Rewrite the answer to the previous question using it. Hint: make sure the variable names are in "".
    
10A. The `one_of()` function selects column names from the whole data set.
```{r}
select(vehicles, one_of("cyl", "displ", "hwy", "cty"))
```



11. What happens if you include the name of a variable multiple times in
    a `select()` call?
    
11A. dplyr will ignore a repeated variable parameter and only show the variable in the first position it was called (left to right)
```{r}
select(vehicles, cty, cyl, displ, hwy, cty, cty)
```


 
12. Make a new variable hwygpm where hwygpm=gallons per mile on the highway=1/hwy

12A.
```{r}
vehicles2 = mutate(vehicles, hwygpm = 1/hwy)
vehicles2
```


13. Make a new variable ctykpl where ctykpl=kilometers per liter in the city=0.42517*cty

13A. 
```{r}
vehicles2 = mutate(vehicles2, ctykpl = 0.42517*cty)
vehicles2
```


14. Make a new variable that is the dfference between hwy and cty.

14A. 
```{r}
vehicles2 = mutate(vehicles2, mpgDif = hwy-cty)
vehicles2
```


15. How many vehicles have a negative difference between hwy and cty? Any common features of some of the vehicles?

15A. Vehicles with negative hwy/cty differentials tend to be electric vehicles which are less efficient in the stop-and-go of cities.
```{r}
arrange(vehicles2, mpgDif)
```

